tl19-07_manuscript_figures

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_manuscript_figures.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

---
title: "TL19-07_Audrey_24h_chabaudi"
date: "17/06/2020"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

@@

To take into account:

- SBP is more synchronous than MT, do we have counts?

- MT and SBP may seem very different due to differences in the amount of material for sequencing.

Suggestions:

- PCA of SBP on its own, and MT on its own, UMAP maybe too?

- Use DESeq2 to get the diff expr genes between time points and then use this to plot a PCA.

- z-score plots: Have a max/min expression heatmap instead of showing the median logTPM.

- Just remove the Unknown sample since it contributes nothing without an identity.

- The value of the expression threshold (variable: min_median) is very high (over 4 TPM) - is this reasonable?

Top_n_genes:

~~- Remove Unknown entirely by re-saving top_genes table. Also need to separate the top_n bit using rMT and SBP.~~

~~- Why is there a NULL printed right at the end? Issue with names~~

Questions:

- What day are these samples taken? d6?

- How many passages had 'SBP' been through?

- What are the differential counts of the samples taken? - Audrey's are on the Shared folder.

# Audrey's experimental description

24 Hours cycle P. chabaudi RNA seq.

Aim: Determine pir expression for each parasite stage in recently mosquito transmitted P. chabaudi parasites.

Principle: using a stabilate from recently mosquito transmitted P. chabaudi parasites infected C57BL/6 and extract RNA at each stage of parasite cycle (every 3 hours) during 24h.

Animals: 4 C57BL/6 J CD45.1 males for stabilates thawing (2 in RL, 2 in NL) and 27 C57BL/6 J CD45.1 females for experiments (12 in NL and 15 in RL) (issued 17/08/2016).

* AV16-35	2 C57BL/6 J CD45.1 males		RL.

* AV16-36	2 C57BL/6 J CD45.1 males		NL.

* AV16-37	15 C57BL/6 J CD45.1 females	RL.

* AV16-38	12 C57BL/6 J CD45.1 females	RL.

Experiment:

* d-14	11am	 Infected 2 mice (males) with recently mosquito transmitted parasite stabilate (MTC tower 6 box 6.1 position 31-65 (late troph)) in NL (AV16-36).

* d-14	5pm	Infecte 2 mice (males) with recently mosquito transmitted parasite stabilate (MTC tower 6 box 6.1 position 31-65 (late troph)) in RL (AV16-35).

* d-7	11am	pass AV16-35 into 15 mice (females) 105 iRBC/ mouse IP (AV16-37) RL.

* d0	 2pm	smear 3 mice and bleeding .

* d0	 2-5pm 	pellet parasites and trizol of 2pm samples.

* d0	 5pm	smear 3 mice and bleeding .

* d0	 5-8pm 	pellet parasites and trizol 5pm samples.

* d0	 8pm 	smear 3 mice and bleeding.

* d0	 8-11pm 	pellet parasites and trizol 8pm samples.

* d0	 11pm	smear 3 mice and bleeding .

* d0	11pm-2am 	pellet parasites and trizol 11pm samples.

* d1	 2am	smear 3 mice and bleeding .

* d1	 2-5am 	pellet parasites and trizol of 2am samples.

* d1	 5am	smear 3 mice and bleeding .

* d1	 5-8am 	pellet parasites and trizol 5pm samples.

* d1	 8am 	smear 3 mice in each group and bleeding .

* d1	 8-11am 	pellet parasites and trizol 8pm samples.

* d1	 11am	smear 3 mice in each group and bleeding .

* d1	 11am 	bleeding 3 mice (AV16-37) and parasite culture.

* d1	11am-2pm 	pellet parasites and trizol 11pm samples.

* d1	 4-6pm	schizonte pellet and trizol.

Dates:

d-14 : 25/08/2016.

d-7 : 01/09/2016.

d0 : 08/09/2016.

d1 : 09/09/2016.

@@

```{r global_save_var}
#Make this TRUE if you want to run this script and save all the plots.
global_save_var <- TRUE
```

```{r tpm, results = 'hide', message = FALSE}
# # Calculating the transcript-per-million 
# 
# We need to write a function to calculate the TPM from the count data we have generated. This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.

tpm = function(count_data, lengths = chabaudi_transcript_lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
    }
  } else {
    stop('Error: Count data and/or length table supplied do not have an id column.')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
  } else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  lengths.table = matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r agree function}

agree = function(input_list){
  for(i in 1:(length(input_list)-1)){
    input_list[[i+1]] = input_list[[i]][ input_list[[i]] %in% input_list[[i+1]] ]
  }
  return(input_list[[length(input_list)]])
}

disagree = function(input_list){
  for(i in 1:(length(input_list)-1)){
    input_list[[i+1]] = input_list[[i]][ !input_list[[i]] %in% input_list[[i+1]] ]
  }
  return(input_list[[length(input_list)]])
}

mround <- function(x,base){ 
  base*round(x/base) 
} 
```

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}
```

```{r pca_function}
pca_function <- function(input_ct, 
                         genevec = NULL, 
                         scale.pca = TRUE, 
                         three.dimensional = FALSE,
                         legend_settings = list(),
                         colornames = NULL){

  twentyfour_tpm_pca <- input_ct$tpm
  if(!is.null(genevec)){
    twentyfour_tpm_pca <- input_ct$tpm %>% filter(Geneid %in% genevec)

    ### Re-scaling the TPM data
    twentyfour_tpm_pca_rescale <- lapply(colnames(twentyfour_tpm_pca), 
                                         function(column_name){
                                           column <- getElement(twentyfour_tpm_pca, 
                                                                paste(column_name))
                                           if(is.numeric(column)){
                                             #is.numeric to avoid trying to scale the Geneid column
                                             tpm_scale_factor <- 1e6 / sum(column)
                                             tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                           }else{
                                             tpm_scaled <- column
                                           }
                                           return(tpm_scaled)
                                         }) %>% 
      bind_cols

    colnames(twentyfour_tpm_pca_rescale) <- colnames(twentyfour_tpm_pca)
    twentyfour_tpm_pca <- twentyfour_tpm_pca_rescale
    ###

  }
  rownames(twentyfour_tpm_pca) <- twentyfour_tpm_pca$Geneid
  twentyfour_tpm_pca <- select(twentyfour_tpm_pca, -Geneid) %>% 
    prcomp(center = TRUE, scale. = scale.pca)
  #Need to look into what exactly scale and center are doing
  twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
  twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

  if(is.null(colornames)){
    # colornames <- c(cbPalette, 'black', 'red')
    colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
    colornames <- setNames(colornames, 
                           sort(unique(twentyfour_tpm_pca_coords$time)))
  }

  if(three.dimensional){
    pl <- plot_ly(twentyfour_tpm_pca_coords, 
                  x  = ~PC1, y = ~PC2, z = ~PC3,  mode = 'markers',
                  type = 'scatter3d', color = ~time, text = ~sample,
                  colors = colornames,
                  # size = ~total_features,
                  marker = list(opacity = 1, size = 5),
                  symbol = ~transmission, 
                  symbols = c('circle','x')) %>% 
      layout(scene = list(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
                          yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
                          zaxis = list(title = paste0('PC3 ',pc3_importance,'%'))))
  }else{
    pl <- plot_ly(twentyfour_tpm_pca_coords, 
                  x  = ~PC1, y = ~PC2,
                  type = 'scatter', mode = 'markers',
                  color = ~time, text = ~sample,
                  colors = colornames,
                  # size = ~total_features,
                  marker = list(opacity = 1, size = 16),
                  symbol = ~transmission, symbols = c('circle','x')) %>% 
      layout(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
             yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
             legend = legend_settings)
  }
  pl
}
```

```{r function_draw_heatmap_chabaudi_24h}
# Function to draw a heatmap with a list of genes

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")

# Note that there may seem to be discrepencies between the data and the heatmap because this function may be using the median and/or mean value of the TPM
draw_heatmap_chabaudi_24h <- function(genelist, 
                                scale = 'max', 
                                threshold = 0, 
                                datasets = twentyfour_ct, 
                                filter_tpm = NULL,
                                stage_diff = NULL, 
                                filter_stage = NULL, 
                                title_name = NULL, 
                                drop_stage = NULL, 
                                log_it = TRUE,
                                collate_samples_by_stage = TRUE,
                                order_by_med_tpm = TRUE,
                                mean_or_median = 'mean',
                                gene_names_for_title = NULL,
                                ...){
  #genelist <- cir_id
  #For de-bugging:

  # scale = 'max';
  # threshold = 0;
  # datasets = twentyfour_ct;
  # filter_tpm = NULL;
  # stage_diff = NULL;
  # filter_stage = NULL;
  # title_name = NULL;
  # drop_stage = NULL;
  # log_it = TRUE;
  # collate_samples_by_stage = TRUE;
  # order_by_med_tpm = TRUE;
  # mean_or_median = 'mean';

  # if(!is.null(datasets$tpm)){
  #   #Probably means only one dataset was input so lapply will malfunction
  #   datasets <- list(tpm = datasets)
  # }

  if(log_it == TRUE){
    threshold <- log1p(threshold)
    title <- 'log(TPM + 1)'
  }else{
    title <- 'TPM'
  }

  expt_tpms_top_genes = lapply(datasets, function(tpm_input){
    # tpm_input <- datasets[[1]]
    tpms = tpm_input$tpm %>% 
      #Melt the data frame for using dplyr
      filter(Geneid %in% genelist) %>% 
      melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm')

    if(collate_samples_by_stage){ 
      #Create a new variable 'stage' from the sample names without rep numbers
      tpms <- mutate(
        tpms, 
        stage = str_extract(
          sample,
          pattern = '.+(?=_[[:digit:]]{1,2})'
        )
      )
    }else{
      tpms <- mutate(tpms, 
                     stage = sample)
    }

    tpms <- mutate(tpms, time = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
      #Convert the times given in 'h' to robustly coded times in 24h clock
      mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
                           strptime(time, format="%Hh") %>% format(., "%Hh"),
                           time)) %>% 
      mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
                           strptime(time, format="%I%p") %>% format(., "%Hh"),
                           time))

    colnam = sort(unique(tpms$stage))

    if(mean_or_median == 'median'){
      tpms = tpms %>% 
        group_by(Geneid, stage, time) %>%  
        dplyr::summarise(mean_tpm = median(tpm)) %>%  
        group_by(stage) %>% 
        recast(Geneid ~ stage + time + 'mean_tpm', median) #recast into short-form for complex heatmap
    } else if (mean_or_median == 'mean'){
      tpms = tpms %>% 
        group_by(Geneid, stage, time) %>%  
        dplyr::summarise(mean_tpm = mean(tpm)) %>%  
        group_by(stage) %>% 
        recast(Geneid ~ stage + time + 'median_tpm', mean) #recast into short-form for complex heatmap
    }

    colnames(tpms) = c('Geneid',as.character(colnam))

    if(!is.null(order_vector)){
      tpms <- tpms[,
                   ordering(colnames(tpms),
                            c('Geneid',
                              order_vector))]
      #Ordering using the ordering function and order_vector defined in the functions section.
    }

    #Apply drop_stage
    if(!is.null(drop_stage)){
      tpms <- select(tpms, -!!(paste(drop_stage)))
    }
    #Apply stage_diff to get genes which are expressed higher in one stage than another
    if(!is.null(stage_diff)){
      if(is.na(stage_diff[1])|is.na(stage_diff[2])){
        stop('stage_diff must be a vector of two named stages to select genes with TPM greater in the former than the latter')
      }
      tpms <- tpms[select(tpms, paste(stage_diff[1])) > select(tpms, paste(stage_diff[2])), ]
    }
    #Apply filter_tpm
    if(!is.null(filter_tpm)){
      tpms <- filter_at(tpms, vars(-Geneid), any_vars(. > filter_tpm))
    }
    #Apply filter_stage to remove genes expressed below filter_tpm in that stage
    if(!any(is.null(filter_stage), is.null(filter_tpm))){
      tpms <- filter(tpms, paste(filter_stage) >= filter_tpm)
    }
    #Remove PCHAS_ from the gene names for simplicity
    # rownames(tpms) = gsub(as.character(tpms$Geneid), pattern = 'PCHAS_', replacement = '')
    #Use the protein names for ease of understanding
    rownames(tpms) = chabaudi_gene_info$handy_descriptions[match(tpms$Geneid, 
                                                           chabaudi_gene_info$Geneid)]
    tpms = select(tpms, -Geneid)

    if(log_it == TRUE){
      tpms <- log1p(tpms)
    }

    if(tpm_input$experiment %>%  is.null){
      tpm_input$experiment = ''
    }
    list = list(tpms, tpm_input$experiment)
    names(list) = c('tpm','experiment')
    return(list)
  })

  if(
    !all(
      unlist(
        lapply(
          expt_tpms_top_genes, 
          function(x) all(
            x$tpm < threshold
          )
        )
      )
    )
  ){
    if(scale %in% 'max'){
      maxcol <- my_max(unlist(lapply(expt_tpms_top_genes, 
                                     function(tpms){my_max(tpms$tpm)})))
    }else{
      maxcol = scale
    }
  }

  if(is.na(maxcol)|maxcol<1){maxcol <- 1}

  if(order_by_med_tpm){
    #Work out the order by median TPM
    medians_across_expt_mat <- as.matrix(
      sapply(
        lapply(
          expt_tpms_top_genes, 
          function(x){
            rowMedians(as.matrix(x$tpm),
                       na.rm = TRUE) 
          }
        ), 
        rbind
      )
    )
    tpm_order <- order(
      rowMedians(
        medians_across_expt_mat, 
        na.rm = TRUE),
      decreasing = TRUE
    )
  }

  #Applying the threshold last because NAs mess things up
  expt_tpms_top_genes <- lapply(expt_tpms_top_genes, function(expt){
    tpms <- expt$tpm
    tpms = apply(tpms, 2, function(x){
      ifelse(
        x >= threshold, 
        x, 
        NA
      )
    }
    ) 
    #Get NA for values under threshold
    list = list(tpms, expt$experiment)
    names(list) = c('tpm','experiment')
    return(list)
  })

  heatmaps = lapply(expt_tpms_top_genes, function(tpms){
    if(tpms$experiment != last(expt_tpms_top_genes)$experiment){
      #If it is the final element don't remove the row names
      rownames(tpms$tpm) = c()
    }
    #Setting the name if one is not provided
    if(is.null(title_name)){
      title_name = tpms$experiment
    }

    #Re-order both data.frames by the median highest expressed pirs
    if(order_by_med_tpm){
      tpms$tpm <- tpms$tpm[tpm_order,, drop = FALSE]
    }

    # print(nrow(tpms$tpm))
    #Drawing each heatmap
    ht = Heatmap(tpms$tpm, 
                 name = 'TPM', 
                 col = colorRamp2(c(threshold, 
                                    maxcol/4, 
                                    maxcol/2, 
                                    3*maxcol/4, 
                                    maxcol), 
                                  viridis(5)),
                 cluster_rows = FALSE, 
                 cluster_columns = FALSE,
                 column_title = paste(title_name),
                 column_title_gp = gpar(fontsize = 7),
                 column_names_gp = gpar(fontsize = 7),
                 row_names_gp = gpar(fontsize = ifelse(
                   nrow(tpms$tpm) > 13, 
                   12*(13/nrow(tpms$tpm)), 
                   12)
                 ),
                 na_col = 'white', border = TRUE,
                 heatmap_legend_param = list(
                   title = paste(title), 
                   at = round(
                     c(threshold, 
                       maxcol/4, 
                       maxcol/2, 
                       3*maxcol/4, 
                       maxcol), 
                     digits = 1)
                 )
                 # ,...
    )
  })

  ht_list = Reduce('+', heatmaps) #To add the list elements together
  draw(ht_list, 
       row_title_side = 'left',
       column_title = bquote(~italic('P. chabaudi')~'24h asexual life cycle stage - '~italic(.(gene_names_for_title))~'genes'), 
       column_title_side = 'bottom',
       heatmap_legend_side = 'left')
}
```

```{r plotly_biplot_function}
plot_ly_biplot <- function(pca_object = twentyfour_tpm_pca, 
                           prin_components = c(1,2,3),
                           fontsize = 7,
                           scale.loads = 600,
                           separate_multigene = FALSE,
                           three_dimensions = FALSE,
                           col_var = 'time',
                           gene_info = chabaudi_gene_info,
                           textposition = "top right"){

  #Check that three components are provided for three dimensional biplot
  if(three_dimensions){
    if(length(prin_components) != 3){
      stop('For three dimensions, three principal components must be given for prin_components argument')
    }
  }

  pca_object_rotations <- rownames_to_column(data.frame(pca_object$rotation), 
                                             var = 'sample')  %>% 
    dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
    dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
    dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

  #Setting the colours, which come from each life cycle experiment, called 'stage'.

  colornames <- setNames(
    viridis(
      length(
        unique(
          getElement(
            pca_object_rotations, 
            paste(col_var)
          )))),
    sort(
      unique(
        getElement(
          pca_object_rotations, 
          paste(col_var)
        ))))

  pca_object_loadings <- data.frame(Geneid = rownames(pca_object$x),
                                    pca_object$x)

  pca_object_loadings$Geneid <- paste0(
    str_replace(
          as.character(pca_object_loadings$Geneid),
          pattern = ' ',
          replacement = ''),
    ' ',
    gene_info$handy_descriptions
    [
      match(
        c(str_replace(
          as.character(pca_object_loadings$Geneid),
          pattern = ' ',
          replacement = '')
        ),
        gene_info$Geneid
      )
      ]
  )

  if(separate_multigene){
    #If the gene is a fam or a fam gene we want to know - haven't excluded pseudogenes here
    pca_object_loadings$multigene <- ifelse(
      #if it doesn't have fam or fam-something in its name then just say gene
    is.na(
      str_extract(
        pca_object_loadings$Geneid,
        pattern = 'fam-.|fam'
      )
    ),
    'Gene loadings',
    paste0(
      str_extract(
        pca_object_loadings$Geneid,
        pattern = 'fam-.|fam'
      ),
      ' gene loadings'
    )
  )
}else{
  pca_object_loadings$multigene <- rep('Gene loadings',
                                       stages = nrow(pca_object_loadings))
}

#Degrees of variation for the axes
pca_importance <- as.data.frame(summary(pca_object)$importance)
pc_importance_1 <- round(pca_importance[2,prin_components[1]]*100, digits = 1)
pc_importance_2 <- round(pca_importance[2,prin_components[2]]*100, digits = 1)
if(three_dimensions){
  pc_importance_3 <- round(pca_importance[2,prin_components[3]]*100, digits = 1)
}

#scale.loads controls how long the lines are, the longer they are the more prominent the loadings but the less
# obvious the genes.

pl <- plot_ly()

#For the font of the labels
t <- list(
  size = fontsize
  )

#For loop to add the PCA lines of the samples
for(k in 1:nrow(pca_object_rotations)) {
  #Have 0 so that the line comes from the origin (0,0)
  #PC first
  x <- c(0, pca_object_rotations[k,(prin_components[1]+1)])*scale.loads
  #PC second
  y <- c(0, pca_object_rotations[k,(prin_components[2]+1)])*scale.loads

  if(three_dimensions){
    #PC third
    z <- c(0, pca_object_rotations[k,(prin_components[3]+1)])*scale.loads
    df_plotly <- data.frame(x = x, 
                            y = y, 
                            z=z,
                            stage = getElement(pca_object_rotations, 
                                               paste(col_var))[k],
                            sample = pca_object_rotations$sample[k])
  }else{
    df_plotly <- data.frame(x = x, 
                            y = y,
                            stage = getElement(pca_object_rotations, 
                                               paste(col_var))[k],
                            sample = pca_object_rotations$sample[k])
  }

  #colour based on sample
  color <- colornames[
    match(
      getElement(
        pca_object_rotations, 
        paste(col_var)
      )[k], 
      rev(names(colornames)
      )
    )
    ]

  #Add these to the plot
  #To avoid duplication in the legend don't show it if the name is duplicated
  if(duplicated(getElement(pca_object_rotations, paste(col_var)))[k]){
    showlegend <- FALSE
  }else{
    showlegend <- TRUE
  }

  #Add the lines for the samples depending on their position in the PCA
  if(three_dimensions){
    pl <- pl %>%
      add_trace(data = df_plotly,
                x=~x,
                y=~y,
                z=~z,
                type='scatter3d',
                mode='line',
                color = ~stage,
                colors = colornames,
                line = list(width = 2),
                text = pca_object_rotations[k,1],
                name = getElement(pca_object_rotations, paste(col_var))[k],
                opacity = 1,
                showlegend = showlegend,
                legendgroup = getElement(pca_object_rotations, paste(col_var))[k],
                #Note that 3d plots only accept a few shapes e.g. hexagon is not accepted
                symbols = c('circle', 
                            'circle-open',
                            'square',
                            'square-open',
                            'x',
                            'diamond')) %>% 
            add_text(x = df_plotly$x[2],
               y = df_plotly$y[2],
               z = df_plotly$z[2],
               text = df_plotly$sample, 
               textfont = t,
               textposition = textposition,
               showlegend = ifelse(k == 1, TRUE, FALSE), #Only want this option to be TRUE once
               legendgroup = 'Toggle text',
               name = 'Toggle text')
  }else{
    pl <- pl %>% 
      add_trace(data = df_plotly,
                x=~x,
                y=~y,
                type='scatter',
                mode='line',
                color = ~stage,
                colors = colornames,
                line = list(width = 2),
                text = pca_object_rotations[k,1],
                name = getElement(pca_object_rotations, 
                                  paste(col_var))[k],
                opacity = 1,
                showlegend = showlegend,
                legendgroup = getElement(pca_object_rotations, 
                                         paste(col_var))[k],
                symbols = c('circle', 
                            'circle-open',
                            'square',
                            'square-open',
                            'x',
                            'diamond')) %>% 
      #I don't know why it only works if we put the symbols argument here and not in the add_trace later??
      add_text(x = df_plotly$x[2],
               y = df_plotly$y[2],
               text = df_plotly$sample,
               textfont = t,
               textposition = textposition,
               showlegend = ifelse(k == 1, TRUE, FALSE), #Only want this option to be TRUE once
               legendgroup = 'Toggle text',
               name = 'Toggle text')
  }
}

#Adding the gene loadings

if(three_dimensions){
  #Make data frame of the gene loadings
  pca_object_loadings_df <- select(data.frame(pca_object_loadings), c(Geneid, 
                                                                      multigene, 
                                                                      paste0('PC', prin_components[1]),
                                                                      paste0('PC', prin_components[2]),
                                                                      paste0('PC', prin_components[3])))
  colnames(pca_object_loadings_df) <- c('Geneid',
                                        'multigene',
                                        'prin_com_1',
                                        'prin_com_2',
                                        'prin_com_3')
  pl <- pl %>% 
    add_trace(data = pca_object_loadings_df, 
              inherit = FALSE,
              x = ~prin_com_1, 
              y = ~prin_com_2, 
              z = ~prin_com_3,
              type = 'scatter3d', 
              mode = 'markers',
              marker = list(color = 'black',
                            size = 4,
                            opacity = 0.7),
              text = ~Geneid,
              name = ~multigene,
              symbol = ~multigene,
              showlegend = TRUE,
              legendgroup = ~multigene) %>% 
    layout(
      scene = list(
        xaxis = list(
          title = paste0(
            'PC', 
            prin_components[1], 
            ' ', 
            pc_importance_1,
            '%'
          )
        ),
        yaxis = list(
          title = paste0(
            'PC', 
            prin_components[2], 
            ' ',
            pc_importance_2,
            '%'
          )
        ),
        zaxis = list(
          title = paste0(
            'PC', 
            prin_components[3], 
            ' ',
            pc_importance_3,
            '%'
          )
        )
      )
    )
    }else{
      #Make data frame of the gene loadings
      pca_object_loadings_df <- select(data.frame(pca_object_loadings), 
                                       c(Geneid, 
                                         multigene, 
                                         paste0('PC', prin_components[1]),
                                         paste0('PC', prin_components[2])))
      colnames(pca_object_loadings_df) <- c('Geneid',
                                            'multigene',
                                            'prin_com_1',
                                            'prin_com_2')
      pl <- pl %>% 
        add_trace(data = pca_object_loadings_df, 
                  inherit = FALSE,
                  x = ~prin_com_1, 
                  y = ~prin_com_2,
                  type = 'scatter', 
                  mode = 'markers',
                  marker = list(color = 'black',
                                size = 6,
                                opacity = 0.9),
                  text = ~Geneid,
                  name = ~multigene,
                  symbol = ~multigene,
                  showlegend = TRUE,
                  legendgroup = ~multigene) %>% 
        layout(
          xaxis = list(
            title = paste0(
              'PC', 
              prin_components[1], 
              ' ', 
              pc_importance_1,
              '%'
            )
          ),
          yaxis = list(
            title = paste0(
              'PC', 
              prin_components[2], 
              ' ',
              pc_importance_2,
              '%'
            )
          )
        )
    }
pl
}

#For debugging
# k <- 6

#Switching around the add_text and add_trace for the 'Toggle text' (so that it goes first) seems to bugger up the colours
```

@@

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r function_firstup}
#Capitalize first letter
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
```

```{r function_density_samples}
density_samples = function(input, save = FALSE){
  #Get some colours for the lines
  # colours = rainbow_hcl(length(colnames(input$tpm)[-1]))
  colours = rep_len(cbPalette, length.out = length(colnames(input$tpm)[-1]))
  list_dens = lapply(colnames(input$tpm)[-1], function(sample){
    data = as.numeric(unlist(select(input$tpm, paste(sample))))
    dens = density(log1p(data), bw = bw_mean)
    #Use the index of the column to get the colour
    colouring = colours[grep(colnames(input$tpm)[-1], pattern = paste(sample))]
    list_return = list(dens, colouring)
    names(list_return) = c('dens','colouring')
    list_return
  })
  max_dens = max(unlist(lapply(list_dens, function(x){max(x$dens$y)})))
  max_x = max(unlist(lapply(list_dens, function(x){max(x$dens$x)})))
  line_width = 1
  par(new=FALSE)
  if(save){
    pdf(file = paste0(input$experiment, '_tl19-07_densityplot_fixedbw.pdf'), 
        width = 15/2.54, 
        height = 11.5/2.54)
  }
  lapply(seq_along(list_dens), function(index){
    x = list_dens[[index]]
    if(index <= 9){
      plot(x$dens, 
           xlim = c(0, max_x), 
           ylim = c(0,max_dens), 
           type = 'l', 
           col = x$colouring,
           main = paste(input$experiment),
           lwd = line_width)
    }else if(index <= 18){
      plot(x$dens, 
           xlim = c(0, max_x), 
           ylim = c(0,max_dens), 
           type = 'l', 
           col = x$colouring,
           main = paste(input$experiment),
           lwd = line_width,
           lty = 2)
    } else {
      plot(x$dens, 
           xlim = c(0, max_x), 
           ylim = c(0,max_dens), 
           type = 'l', 
           col = x$colouring,
           main = paste(input$experiment),
           lwd = line_width,
           lty = 3)
    }
    abline(v = log1p(min_median))
    legend(max_x-4, max_dens, 
           legend = paste(colnames(input$tpm)[-1]), 
           col = colours,
           lty = c(rep(1, times = 9), rep(2, times = 9), rep(3, times = 9)),
           cex = 0.5)
    par(new=TRUE)})
  if(save){
    dev.off()
  }
}
```

```{r function_total_tpm_expression}
total_tpm_expression <- function(list_genes, 
                                 save = FALSE,
                                 tpm_data = twentyfour_ct,
                                 stage_str_extract_regex_pattern = '(?<=_).+(?=_[[:digit:]]+)|sch.culture'){
  lapply(names(list_genes), 
         function(list_name){
           paste(list_name)
           gene_id <- getElement(list_genes, paste(list_name))
           list_tpms_line <- lapply(tpm_data, 
                                    function(input){
             tpm_data = input$tpm %>% 
               filter(Geneid %in% gene_id) %>% 
               melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
               mutate(stage = str_extract(sample, 
                                          pattern = paste(stage_str_extract_regex_pattern))) %>% 
               group_by(stage, sample) %>% 
               dplyr::summarise(total_tpm = sum(tpm))

             #All genes detected
             tpm_data_all_genes <- input$tpm %>% 
               melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
               mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),
                                            `[`,
                                            1))) %>% 
               # filter(tpm > threshold) %>% 
               group_by(stage, sample) %>% 
               dplyr::summarise(total_tpm = n()) %>% 
               ungroup %>%  
               group_by(stage) %>% 
               dplyr::summarise(total_tpm_median = median(total_tpm),
                                Number_of_replicates = n(),
                                total_tpm_max = max(total_tpm),
                                total_tpm_min = min(total_tpm)) %>% 
               arrange(stage)

             #Make stages with no cirs in the sample give a zero 
             if(!all(colnames(input$tpm) %in% tpm_data$sample)){
               missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
               missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
                 data.frame(stage = unlist(str_extract(missing_st, 
                                                       pattern = paste(stage_str_extract_regex_pattern))),
                            sample = missing_st,
                            total_tpm = 0)
               }))
               tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
             }

             #Add experiment column
             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
             max_tpm = if(max(tpm_data$total_tpm) < 10){
               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
             }else{
               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
             }

             #Get the average data for the line graph
             tpm_data_line = tpm_data %>% 
               ungroup %>% 
               group_by(stage, experiment) %>% 
               dplyr::summarise(total_tpm_median = median(total_tpm),
                                Number_of_replicates = n(),
                                total_tpm_max = max(total_tpm),
                                total_tpm_min = min(total_tpm)) %>% 
               arrange(stage)

             tpm_data_line$total_tpm_median_allgenes <- tpm_data_all_genes$total_tpm_median

             return(tpm_data_line)
           })

>>>>

---
title: "tl19-07_manuscript_figures"
author: "Timothy Little"
date: "06/11/2020"
output: html_document
---

@@

@@

@@

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```
